ms_figures

### 2018-01-25 analysis/disturbance/analysis_disturbance_chronos.Rmd
update figures

@@

>>>>

```{r}
gc_trees50<- c()

  au <- df %>% computeGCtrees(treshold = 50) %>% data.frame() %>% mutate(site = paste(d))
  gc_trees50 <- rbind(gc_trees50, au)
}

gc_trees50  <- gc_trees50  %>% 
  mutate(site_sorted = case_when(
    site == "gc_sj" ~ "0_sj", 
    site == "gc_caH" ~ "1_caH",
    site == "gc_caL" ~ "2_caL"))

gc_trees %>%   
  write.csv(file=paste(di, "data/disturbance/disturbance_trees_gc_treshold50.csv", sep=""), row.names = FALSE)
```

@@

```{r}
gc_treesMed50 <- c()

  au <- df %>% computeGCtreesMedian(treshold = 50) %>% data.frame() %>% mutate(site = paste(d))
  gc_treesMed50 <- rbind(gc_treesMed50, au)
}

gc_treesMed50  <- gc_treesMed50  %>% 
  mutate(site_sorted = case_when(
    site == "gc_sj" ~ "0_sj", 
    site == "gc_caH" ~ "1_caH",
    site == "gc_caL" ~ "2_caL"))

gc_treesMed50 %>%   
  write.csv(file=paste(di, "data/disturbance/disturbance_trees_gc_treshold_median50.csv", sep=""), row.names = FALSE)
```

### 2018-01-25 man/ms_figures.Rmd
update rwi climate with selected monts
update figures

---
title: "ms_figures"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

@@

Figure 1. Distribution of *Quercus pyrenaica* forests in Iberian Peninsula (a) and in Sierra Nevada mountain range, where three cluster of oak populations have been identified [@PerezLuque2015] (showed in different colour) (b). A grid of with the MODIS pixels for each population is shown (see material and methods). Detailed location of the sampling sites: northern (San Juan, SJ) (c) and southern ones (Cáñar: CALow and CAHigh)(d). Colour Orthophotography of 2009 from Regional Ministry of the Environment, Regional Government of Andalusia.

@@

Figure 2. Comparison of EVI profile for the reference period (*gray*) and during the 2005 (*green*) and 2012 (*blue*) drought events (a). EVI standardized anomaly ($EVI~sa$) during the period 2000-2016 for northern and southern populations (b). Error bars show standard error.

@@

Figure 3. Response of northern (*black*) and southern (*blue*) populations of *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; a) and tree growth (BAI; b) for the years 2005 and 2012. Different letters above bars indicate significant post hoc differences between groups (see material and methods).

@@

Figure 4. Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods).

@@

```{r, warning=FALSE}
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

cronos_bai <- cro_sites %>% ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.2, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  theme(panel.grid = element_blank(),
        legend.position = c(0.1,.8),
        strip.background = element_blank()) + 
  scale_colour_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_fill_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10)) 
  # scale_y_continuous(sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  # geom_line(aes(x=year, y=n_trees+40, colour=site))

@@

Figure 5. Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA_Low; *pink*) and high-elevation (CA_High, *purple*) sites. Shading areas coorespond to standard error of the mean. Number of series are displayed in the upper plot.

@@

>>>>

---
output:
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: template_ms.docx
bibliography: ../refs/references.bib
csl: ../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

@@

**Figure 1**. Distribution of *Quercus pyrenaica* forests in Iberian Peninsula (a) and in Sierra Nevada mountain range, where three cluster of oak populations have been identified [@PerezLuque2015] (showed in different colour) (b). A grid of with the MODIS pixels for each population is shown (see material and methods). Detailed location of the sampling sites: northern (San Juan, SJ) (c) and southern ones (Cáñar: CALow and CAHigh)(d). Colour Orthophotography of 2009 from Regional Ministry of the Environment, Regional Government of Andalusia.

@@

**Figure 2.** Comparison of EVI profile for the reference period (*gray*) and during the 2005 (*green*) and 2012 (*blue*) drought events (a). EVI standardized anomaly ($EVI~sa$) during the period 2000-2016 for northern and southern populations (b). Error bars show standard error.

@@

**Figure 3.** Response of northern (*black*) and southern (*blue*) populations of *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; a) and tree growth (BAI; b) for the years 2005 and 2012. Different letters above bars indicate significant post hoc differences between groups (see material and methods).

@@

**Figure 4.** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods).

@@

```{r, warning=FALSE, out.extra='angle=90'}
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

cronos_bai <- cro_sites %>% ggplot(aes(x=year, y=bai_mean/100, colour=site)) + 
  theme_bw() + ylab('BAI') + 
  geom_ribbon(aes(ymin = (bai_mean - bai_se)/100,
                  ymax = (bai_mean + bai_se)/100,
                  fill=site), alpha=.2, colour=NA) +
  geom_line() + 
  geom_point(size=.8) + 
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(size=6),
        legend.position = c(0.1,.8),
        strip.background = element_blank()) + 
  scale_colour_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  scale_fill_manual(values = c('#7570b3','#e7298a','#1b9e77')) +
  ylab(expression(paste("BAI (", cm^2, year^-1, ")", sep=''))) +
  scale_x_continuous(breaks = seq(1815, 2020, by=10)) 
  # scale_y_continuous(sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  # geom_line(aes(x=year, y=n_trees+40, colour=site))

@@

**Figure 5.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA_Low; *pink*) and high-elevation (CA_High, *purple*) sites. Shading areas coorespond to standard error of the mean. Number of series are displayed in the upper plot.

@@

**Figure 6.** Residual Tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the three most recent severe drought years (1995, 2005 and 2012).

# Figure 7
```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',')

rtmax <- rtmax %>% mutate(v_name = 'T max')
rtmin <- rtmin %>% mutate(v_name = 'T min')
rsequia <- rsequia %>% mutate(v_name ='SPEI')
rprec <- rprec %>% mutate(v_name = 'Prec')

df <- rbind(rtmax, rtmin, rsequia, rprec)

# cutre solucion para el problema de PREC 
exdf <- data.frame(
  id = c(16:19),
  varname = rep("prec", 4), 
  month = c('dec...FEB', 'MAR...MAY', 'JUN...AUG', 'SEP...NOV'),
  coef = rep(0, 4), 
  significant = rep('FALSE', 4),
  ci_lower = rep(0, 4),
  ci_upper = rep(0, 4), 
  sig = rep("", 4), 
  month_name = c("Winter", "Spring", "Summer", "Autumn"), 
  v_name = rep("Prec", 4))

exdf <- rbind(exdf, exdf, exdf)
exdf$X <- c(46:57)
exdf$site <- c(rep("SJ", 4), rep("caL", 4), rep("caH", 4))

ddf <- rbind(df, exdf)

label_var <- c("Prec" = "(a) Prec", "SPEI" = "(b) SPEI", 
               "T max" = "(c) T max", "T min" = "(d) T min")

clima_ring <- ddf %>% ggplot(aes(x=id, y=coef, group = site)) + 
  geom_hline(yintercept = 0, colour="gray") +
  geom_bar(aes(fill=site), stat="identity", position = "dodge") +
  facet_wrap(~v_name, labeller = as_labeller(label_var)) + 
  geom_text(aes(label = sig, x = id, y = ifelse(coef > 0, (coef + 0.02), (coef - 0.05))), 
            position = position_dodge(.9), size = 6) + 
  theme_bw() + 
  theme(axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.3),
        legend.position = "bottom",
        strip.background = element_rect(fill=NA, colour=NA),
        strip.text.x = element_text(size = 10, face = "bold")) + 
  scale_x_continuous(breaks = unique(ddf$id), labels = unique(ddf$month_name)) +
  scale_y_continuous(limits = c(-0.4, 0.8), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4, 0.6,0.8)) + 
  scale_fill_manual(values = c("#B2DF8A","#1F78B4","#A6CEE3")) + 
  ylab('Correlation')

clima_ring 
```

**Figure 7.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data (precipitation (a), SPEI (b), maximun (c) and minimun (d) temperatures) for northern site (*green* bars), low-elevation southern site (*light blue* bars) and high-elevation shouthern (*dark blue* bars) site. Asteriks indicate significant ($P < 0.05$) correlation coefficients.

# Figure 8

```{r}
gc_todos <- read.csv(here("data/disturbance", "disturbance_gc.csv"), header = TRUE, sep = ',')

# customize plot (corrected scale)
pos_neg_log <- scales::trans_new('signed_log', 
                                 transform = function(x) sign(x)*log(abs(x)),
                                 inverse = function(x) sign(x)*exp(abs(x)))

custom_ylab <- '% GC'
micolor <- c("#838B8B")

# solution for arrange
# http://felixfan.github.io/stacking-plots-same-x/

custom_corrected_gc <- list(
  geom_bar(stat='identity', fill=micolor, colour=micolor),
  theme_bw(), 
  geom_hline(yintercept=c(50, -50), linetype=2, colour='black'),
  theme(panel.grid.minor = element_blank(), 
        strip.background = element_rect(colour='black',fill='white'))
  )

c <- gc_todos %>% 
  filter(site != 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("1_caH" = "Southern High (CaH)", "2_caL" = "Southern Low (CaL)"))) + 
    ylab(custom_ylab) + custom_corrected_gc

d <- gc_todos %>% 
  filter(site == 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  xlim(layer_scales(c)$x$range$range[1],layer_scales(c)$x$range$range[2]) +
  geom_hline(yintercept = 1, colour = 'gray', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour=micolor, width=.3, size=.2) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("0_sj" = "Northern (SJ)"))) + 
  custom_corrected_gc + ylab('') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-1000, -500, -100, 0, 100, 500)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

plot_medians_corregido <- d + c + plot_layout(ncol=1, heights = c(1, 2))
plot_medians_corregido

**Figure 8.** Comparison of median growth change ($GC$) following [@Nowacki1997] for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

# Figure 9

```{r}
gc_treesMed <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median.csv"), header = TRUE, sep = ',')
gc_treesMed50 <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median50.csv"), header = TRUE, sep = ',')

colorpositivo <- c("#104E8B")
colornegativo <- c("#CD3333")

pgc_median <- gc_treesMed %>% ggplot() +
  geom_bar(aes(x=year, y=per_pos), fill=colorpositivo, colour=colorpositivo, stat = "identity", position = "dodge") + 
  geom_bar(aes(x=year, y=(-1)*per_neg), fill=colornegativo, colour=colornegativo, stat = "identity", position = "dodge") +
  facet_wrap(~site_sorted, nrow = 3, labeller = as_labeller(label_site)) + 
  scale_y_continuous(sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  geom_line(aes(x=year, y=total_tree), colour = 'black') + 
  theme_bw() + ylab('% trees') +
  geom_hline(yintercept=c(50, -50), linetype=2, alpha=.4) +
  theme(panel.grid.minor = element_blank(), 
        strip.background = element_rect(colour='black', 
                                        fill='white')) 
pgc_median

**Figure 9**. Percentage of *Q. pyrenaica* trees affected by GC > 25 % by site. *Black* line shows number of trees (rigth-axis).
